Bow shocks, bow waves, and dust waves

### 2017-12-27 papers/Papers1234/sec-dust-bow-intro.tex
Add paragraph on i-front trapping effect on R0

\subsection{Radiative cooling lengths}
\label{sec:radi-cool-lengths}
\newcommand\M{\ensuremath{\mathcal{M}}}
In this section, we calculate whether the radiative cooling is
sufficiently rapid behind the bow shock to allow the formation of a
thin, dense shell.  Since cooling is least efficient at low densities,
we will assume that the wind bow shock regime applies unless otherwise
specified. We label quantities just outside the shock by the subscript
``0'', quantities just inside the shock (after thermalization, but
before any radiative cooling) by the subscript ``1'', and quantities
after the gas has cooled back to the photoionization equilibrium
temperature by the subscript ``2''.  Assuming a ratio of specific
heats, \(\gamma = 5/3\), the relation between the pre-shock and immediate
post-shock quantities is
\begin{align}
  % \M_1 &= \left(\frac{\M_0^2 + 3} {5\M_0^2 - 1}\right)^{1/2} \\
  \label{eq:shock-n-jump}
  \frac{n_1}{n_0} &= \frac{4 \M_0^2} {\M_0^2 + 3} \\
  \label{eq:shock-T-jump}
  \frac{T_1}{T_0} &= \tfrac1{16} \bigl( 5\M_0^2 - 1 \bigr) \bigl( 1 + 3/\M_0^2 \bigr) \\
  \label{eq:shock-v-jump}
  \frac{v_1}{v_0} &= \left(\frac{n_1}{n_0}\right)^{-1} \ ,
\end{align}
where \(\M_0 = v_0 / \sound\).  The cooling length of the post-shock
gas can be written as
\newcommand\cool{\ensuremath{_{\text{cool}}}}
\begin{equation}
  \label{eq:dcool}
  d\cool = \frac{P_1 v_1} { (\gamma - 1) \bigl(  \mathcal{L}_1 - \mathcal{G}_1 \bigr) }\ ,  
\end{equation}
where \(P_1\) is the thermal pressure and \(\mathcal{L}_1\),
\(\mathcal{G}_1\) are the volumetric radiative cooling and heating
rates.  For fully photoionized gas, we have
\(P_1 \approx 2 n_1 k T_1\), \(\mathcal{L}_1 = n_1^2 \Lambda(T_1)\), and
\(\mathcal{G}_1 = n_1^2 \Gamma(T_1)\), where \(\Lambda(T)\) is the cooling
coefficient, which is dominated by metal emission lines that are
excited by electron collisions, and \(\Gamma(T)\) is the heating
coefficient, which is dominated by hydrogen photo-electrons
\citep{Osterbrock:2006a}. The cooling coefficient has a maximum around
\SI{e5}{K}, and for typical ISM abundances can be approximated as
follows:
\begin{align}
  \label{eq:cooling-coefficient}
  \Lambda_{\text{warm}} &= \num{3.3e-24} \, T_4^{2.3} \, \si{erg.cm^{-3}.s^{-1}}\\
  \Lambda_{\text{hot}} &= \num{e-20} \, T_4^{-1}\, \si{erg.cm^{-3}.s^{-1}} \\
  \Lambda &= \left( \Lambda_{\text{warm}}^{-k} +  \Lambda_{\text{hot}}^{-k} \right)^{-1/k}
      \quad \text{with} \quad k = 3 \ ,
\end{align}
which is valid in the range \(0.7 < T_4 < 1000\).  We approximate the heating coefficient as
\begin{equation}
  \label{eq:heating-coefficient}
  \Gamma = \num{1.77e-24} \, T_4^{-1/2} \, \si{erg.cm^{-3}.s^{-1}} \ ,
\end{equation}
where the coefficient is chosen so as to give \(\Gamma = \Lambda\) at
an equilibrium temperature of \(T_4 = 0.8\).

>>>>

In principle, when the ionization front trapping occurs in the bow
wave regime, then the curves for \(R_0\) will be modified in the
region above the red line because all of the ionizing radiation is
trapped in the shell due to gas opacity, which is not included in
equation~\eqref{eq:tau-thin}.  However, this only happens for our
\SI{10}{M_\odot} star, which has a relatively soft spectrum.
Table~\ref{tab:stars} gives the peak wavelength of the stellar
spectrum for this star as \(\lambda_{\text{eff}} = \SI{0.115}{\um}\), which
is significantly larger than the hydrogen ionization threshold at
\SI{0.0912}{\um}, meaning that only a small fraction of the total
stellar luminosity is in the EUV band and affected by the gas opacity.
The effect on \(R_0\) is therefore small.  For the higher mass stars,
\(\lambda_{\text{eff}} < \SI{0.0912}{\um}\), so the majority of the
luminosity is in the EUV band, but in these cases the ionization front
trapping occurs well inside the radiation bow shock zone, where the
dust optical depth is already sufficient to trap all of the radiative
momentum.

\subsection{Radiative cooling lengths}
\label{sec:radi-cool-lengths}
\newcommand\M{\ensuremath{\mathcal{M}}}
In this section, we calculate whether the radiative cooling is
sufficiently rapid behind the bow shock to allow the formation of a
thin, dense shell.  Since cooling is least efficient at low densities,
we will assume that the wind bow shock regime applies unless otherwise
specified. We label quantities just outside the shock by the subscript
``0'', quantities just inside the shock (after thermalization, but
before any radiative cooling) by the subscript ``1'', and quantities
after the gas has cooled back to the photoionization equilibrium
temperature by the subscript ``2''.  Assuming a ratio of specific
heats, \(\gamma = 5/3\), the relation between the pre-shock and immediate
post-shock quantities is
\begin{align}
  % \M_1 &= \left(\frac{\M_0^2 + 3} {5\M_0^2 - 1}\right)^{1/2} \\
  \label{eq:shock-n-jump}
  \frac{n_1}{n_0} &= \frac{4 \M_0^2} {\M_0^2 + 3} \\
  \label{eq:shock-T-jump}
  \frac{T_1}{T_0} &= \tfrac1{16} \bigl( 5\M_0^2 - 1 \bigr) \bigl( 1 + 3/\M_0^2 \bigr) \\
  \label{eq:shock-v-jump}
  \frac{v_1}{v_0} &= \left(\frac{n_1}{n_0}\right)^{-1} \ ,
\end{align}
where \(\M_0 = v_0 / \sound\).  The cooling length of the post-shock
gas can be written as
\newcommand\cool{\ensuremath{_{\text{cool}}}}
\begin{equation}
  \label{eq:dcool}
  d\cool = \frac{3 P_1 v_1} { 2 \bigl(  \mathcal{L}_1 - \mathcal{G}_1 \bigr) }\ ,  
\end{equation}
where \(P_1\) is the thermal pressure and \(\mathcal{L}_1\),
\(\mathcal{G}_1\) are the volumetric radiative cooling and heating
rates.  For fully photoionized gas, we have
\(P_1 \approx 2 n_1 k T_1\), \(\mathcal{L}_1 = n_1^2 \Lambda(T_1)\), and
\(\mathcal{G}_1 = n_1^2 \Gamma(T_1)\), where \(\Lambda(T)\) is the cooling
coefficient, which is dominated by metal emission lines that are
excited by electron collisions, and \(\Gamma(T)\) is the heating
coefficient, which is dominated by hydrogen photo-electrons
\citep{Osterbrock:2006a}. The cooling coefficient has a maximum around
\SI{e5}{K}, and for typical ISM abundances can be approximated as
follows:
\begin{align}
  \label{eq:cooling-coefficient}
  \Lambda_{\text{warm}} &= \num{3.3e-24} \, T_4^{2.3} \, \si{erg.cm^{-3}.s^{-1}}\\
  \Lambda_{\text{hot}} &= \num{e-20} \, T_4^{-1}\, \si{erg.cm^{-3}.s^{-1}} \\
  \Lambda &= \left( \Lambda_{\text{warm}}^{-k} +  \Lambda_{\text{hot}}^{-k} \right)^{-1/k}
      \quad \text{with} \quad k = 3 \ ,
\end{align}
which is valid in the range \(0.7 < T_4 < 1000\).  We approximate the heating coefficient as
\begin{equation}
  \label{eq:heating-coefficient}
  \Gamma = \num{1.77e-24} \, T_4^{-1/2} \, \si{erg.cm^{-3}.s^{-1}} \ ,
\end{equation}
where the coefficient is chosen so as to give \(\Gamma = \Lambda\) at
an equilibrium temperature of \(T_4 = 0.8\).
